---
title: 'Does better internet access promote economic well-being?'
output:
  html_notebook: 
    code_folding: hide
    theme: cerulean
---

```{r, include=FALSE}
# install.packages("sp")
# install.packages("eurostat")
# install.packages("dplyr")
library(sp)
library(eurostat)
library(dplyr)

colours <- c('#ffffcc','#c7e9b4','#7fcdbb','#41b6c4',
             '#1d91c0','#225ea8','#0c2c84')
```

# Intro

Some hypotheses why Internet access in general could be important for
economic growth

*   digital services an increasingly larger part of the economy
*   very innovative sector: higher returns?
*   more efficient production functions

Or why this could not be

*   digital services rarely create new products
*   focus is on better production technologies
*   big allocative and structural effects; costly adjustments


```{r, results='hide'}
broadband <- eurostat::get_eurostat(id = "isoc_r_broad_h")
income    <- eurostat::get_eurostat(id = "tgs00005")
```



```{r, include=FALSE}
get_year <- function(dta, sel) {
  dta %>% 
  dplyr::pull(time) %>% 
  unique() %>% 
  sel() %>% 
  format("%Y")
}
```

## Data

To answer this question, I obtained data about regional income and broadband internet access
from [Eurostat](http://ec.europa.eu/eurostat). The data comprises a yearly panel that runs
from `r get_year(income, min)` to `r get_year(income, max)` for income and 
from `r get_year(broadband, min)` to `r get_year(broadband, max)` for broadband data.

For simplicity, I take a first glance at the most recent data and visualise 
the distribution across European NUTS2 and NUTS3 regions.

```{r, message=FALSE, results='hide', error=TRUE, warning=FALSE}
income_2015 <- income %>% 
  dplyr::filter(time == "2015-01-01") %>% 
  dplyr::select(-c(unit, time))

income_map_data <- income_2015 %>% 
  dplyr::mutate(cat = cut_to_classes(values, 7, style = "quantile")) %>% 
  eurostat::merge_eurostat_geodata(output_class = "spdf")
```

```{r, fig.width=30, fig.height=30}
fig <- sp::spplot(
  main=list(label="Income",cex=3),
  income_map_data, "cat", 
  xlim = c(-22, 34), ylim = c(35, 70),
  col.regions = c("grey", colours)
)
update(fig, par.settings = list(fontsize = list(text = 40)))
```

```{r, message=FALSE, comment=FALSE}
broadband_2014 <- broadband %>% 
  dplyr::filter(time == "2014-01-01",
                unit == "PC_HH",
                nchar(as.character(geo)) >= 3) %>% 
  dplyr::select(-c(unit, time))


bb_map_data <- broadband_2014 %>% 
  dplyr::mutate(cat = cut_to_classes(values, 7, style = "jenks")) %>% 
  eurostat::merge_eurostat_geodata(output_class = "spdf")
```

```{r, fig.width=25, fig.height=25}
fig <- sp::spplot(
  main=list(label="Broadband Access",cex=3),
  bb_map_data, "cat", 
  xlim = c(-22, 34), ylim = c(35, 70),
  col.regions = c("grey", colours)
)
update(fig, par.settings = list(fontsize = list(text = 35)))
```

## The nexus betwen Income and Internet Access

Here is a scatter plot showing the regions for which data was available in 2015
for income and in 2014 for broadband access. There is a clear positive
relationship. However, this could be driven by a multitude of confounders
and the relationship is riddled with reverse causality. The results below
therefore only point to a fairly strong correlation without any causal 
implications.

```{r, warning=FALSE, fig.height=5}
data <- dplyr::inner_join(broadband_2014, income_2015, by = "geo") %>% 
  dplyr::rename(broadband = values.x, income = values.y)

plot(y = data$broadband, data$income,
     main = "The Relation between Internet Access and Income",
     xlab = "Income in 2015",
     ylab = "Broadband access in 2014")
```


```{r, include = TRUE, results="hide"}
fit <- lm(broadband ~ I(income/1000), data = data)
(results <- summary(fit))
```

### Main findings

Running a regression between the two variables, we find that if the income of a
region inceases by 1000 €, the percentage of households
having internet access in that region is expected to increase by 
`r round(results$coefficients[2, 1], 3)`  percentage points.

